---
title: "p8105_hw3_kpp2126"
author: "Kevin P. Patterson"
date: "2022-10-09"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(ggplot2)
```
## Problem 1
```{r load data}
library(p8105.datasets)
data("instacart")
skimr::skim(instacart)
```
| The goal is to do some exploration of this dataset:
|  (1) Write a short description of the dataset
|  (2) Note the size and structure of the data
|  (3) Describe some key variables
|  (4) Give illustrative examples of observation

The `instacart` dataset has `r nrow(instacart)` observations and `r ncol(instacart)` variables. There are 4 character variables that indicate `aisle`, `department`, `product_name`, and `eval_set`. There are 11 numeric variables that specifically id the order, product, user, number, and aisle, while the rest detail higher level characteristics like the order of when the product added to cart, if it was reordered, day of the week ordered, hour of the day ordered, and days since last order.

| Then, do or answer the following (commenting on the results of each):
|  (1a) How many aisles are there?
|  (1b) Which aisles are the most items ordered from?
|  (note) Arrange aisles sensibly, and organize your plot so others can read it.
```{r number of aisles}
#part 1a
distinctaisles <- instacart %>%
  select(aisle_id, aisle) %>% #including/excluding aisle doesn't make any difference in the final count
  distinct

#part 1b
mostorderedaisle <- instacart %>%
  group_by(aisle_id, aisle) %>%
  count %>%
  arrange(desc(n), .by_group = F) #I added the false condition to override the grouping I did earlier so that I can arrange it by descending order without any attention to group
```
There are `r pull(count(distinctaisles))` aisles, and aisle #`r mostorderedaisle[1,1]` (`r mostorderedaisle[1,2]`) is the most ordered aisle with `r mostorderedaisle[1,3]` orders, followed by aisles #`r mostorderedaisle[2,1]` (`r mostorderedaisle[2,2]`), #`r mostorderedaisle[3,1]` (`r mostorderedaisle[3,2]`), and #`r mostorderedaisle[4,1]` (`r mostorderedaisle[4,2]`) with `r mostorderedaisle[2,3]`, `r mostorderedaisle[3,3]`, `r mostorderedaisle[4,3]` orders, respectively.

|  (2) Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than         10000 items ordered.
|  (note) Arrange aisles sensibly, and organize your plot so others can read it.
```{r items ordered}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(count = n,
         aisle = fct_reorder(aisle, count)) %>%
  ggplot(aes(x = count, y = aisle)) +
  geom_col(width = 0.7) +
  labs(title = "Instacart: Number of Items ordered in each Aisle")
```

|  (4) Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r table three most popular}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(desc(n)) %>%
  knitr::kable(col.names = c('Aisle' , 'Product Name', 'Count', 'Rank'))
```

|  (5) Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table). 
```{r table mean hour}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(meanhour = mean(order_hour_of_day)) %>%
  select(order_dow, product_name, meanhour) %>%
  pivot_wider(names_from = order_dow,
              values_from = meanhour) %>%
  knitr::kable(col.names = c('Product Name', 'Sun', 'Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat'),
               digits = 1)
```

## Problem 2

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). In this spreadsheet, variables `activity.*` are the activity counts for each minute of a 24-hour day starting at midnight.

|  (1) Load, tidy, and otherwise wrangle the data
|  (note) Your final dataset should include all originally observed variables and values
|  (note) Have useful variable names
|  (1a) Include a weekday vs weekend variable
|  (1b) Encode data with reasonable variable classes

```{r load tidy wrangle}
accel_df = read_csv('data/accel_data.csv') %>% 
  janitor:: clean_names() %>%
  pivot_longer(activity_1:activity_1440,
               names_to = "activity_minute",
               names_prefix = "activity_",
               values_to = "activity_count") %>%
  mutate(weekend_weekday = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday"),
         activity_minute = as.numeric(activity_minute),
         day = as.factor(day))

#skimr::skim(accel_df)
```
|  (1c) Describe the resulting dataset (e.g. what variables exist, how many observations, etc)

The resulting tidied dastset contains `r nrow(accel_df)` activity counts and `r ncol(accel_df)` variables. Of these 
`r ncol(accel_df)` variables, there is `week` variable that ranges from 1-5 weeks; there's a `day_id` variable that ranges from 1-29 days; and there's a `day` variable that indicates the day of the week with a `weekend_weekday` variable that indicates if it is a weekend or weekday. Lastly, there are `activity_minute` and `activity_count` that indicate the minute of the day and the corresponding activity count recorded at that minute. 

Traditional analyses of accelerometer data focus on the total activity over the day:

Using your tidied dataset, 
|  (2) Aggregate across minutes to create a total activity variable for each day
|  (2a) Create a table showing these totals. Are any trends apparent?

```{r table totals}

```


Accelerometer data allows the inspection activity over the course of the day: 

|  (3) Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. |  (3a) Describe in words any patterns or conclusions you can make based on this graph.

```{r plot 24-hour}

```


## Problem 3

```{r load read data}
library(p8105.datasets)
data("ny_noaa")
```

*The goal is to do some exploration of this dataset.*

|  (1) To that end, write a short description of the dataset:
|  (1a) noting the size and structure of the data
|  (1b) describing some key variables
|  (1c) indicating the extent to which missing data is an issue

Then, do or answer the following (commenting on the results of each):

|  (2) Do some data cleaning
|  (2a) Create separate variables for year, month, and day. 
|  (2b) Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 
|  (2c) For snowfall, what are the most commonly observed values? Why?

|  (3) Make a two-panel plot showing the average max temperature in January and in July in each station across years. 
|  (3a) Is there any observable / interpretable structure? Any outliers?

|  (4) Make a two-panel plot showing 
|  (4a) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option);
|  (4b) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.


